Women and Highly Educated Toronto Residents Most Concerned About Local Impacts of Climate Change

### 2021-04-27 outputs/paper/paper.Rmd
Paper and scripts update

---
title: "Climate Perceptions in Toronto"
author: "Lorena Almaraz De La Garza"
date: "April 19th, 2021"
keywords: "X, Y, Z"
output:
  bookdown::pdf_document2: default
subtitle: "with data from Open Data Toronto"
toc: FALSE
thanks: 'Code and data are available in this GitHub repository: [lalmaraz/toronto_climate_perceptions](https://github.com/lalmaraz/toronto_climate_perceptions).'
bibliography: references.bib
abstract: "The abstract text occupies this beautiful space right here. \\par
 \\textbf{Keywords:} search, stackoverflow, find, workaround"

@@

```{r libs, include = FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Load libraries
library(bookdown)
library(ggmap)
library(gt)
library(here)
library(httr)
library(janitor)
library(kableExtra)
library(knitr)
library(opendatatoronto)
library(readxl)
library(rio)
library(tidyverse)
```

# Introduction
[@R]

# Data

## Methodology
## Results

# Discussion

## Findings
## Implications
## Limitations
## Future Work

```{r echo=FALSE, message=FALSE, warning=TRUE}

>>>>

---
title: "Women and Highly Educated Toronto Residents Most Concerned About Local Impacts of Climate Change"
author: "Lorena Almaraz De La Garza"
date: "April 26th, 2021"
output:
  bookdown::pdf_document2: default
  fig_caption: yes
subtitle: "A study on climate change perceptions with the City of Toronto's Open Data"
toc: TRUE
thanks: 'Code and data are available in this GitHub repository: [lalmaraz/toronto_climate_perceptions](https://github.com/lalmaraz/toronto_climate_perceptions).'
bibliography: references.bib
abstract: "This report presents a secondary analysis of the City of Toronto's data on residents' perceptions about climate change. The results indicate that respondents' gender, education, and income, affect their self-reported level of concern, and level of information, about climate change. The analysis consists of logistic regression performed with the statistical programming language R. The results contribute to our understanding of the issue at a local level, enabling effective action to address the ongoing threat of climate change.
 \\par
 \\textbf {Keywords:} climate change, climate perceptions, climate concerns, climate action, age and climate, gender and climate, toronto residents, open data toronto"

@@

```{r libraries, include = FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Load libraries
library(bookdown)
library(ggmap)
library(here)
library(httr)
library(janitor)
library(kableExtra)
library(knitr)
library(opendatatoronto)
library(readxl)
library(tidyverse)
library(RCurl)
library(xml2) 
library(rvest) 
library(osmdata) 
library(sf) 
library(performance)
library(modelsummary)
library(patchwork)
library(extrafont)
# Libraries below this comment need to be added to references.bib and cited.

# Get font
font_import()
loadfonts()
```

# Introduction
Although scientific research on the impacts of climate change has been conducted for decades, the last few years have seen a surge in public participation and activism. In particular, youth climate leaders, with unwavering motivation and forthright campaigns, have guided and inspired thousands around the world [@unicef]. In their demands, they often ask that adults enact intergenerational equity, which states that every generation, present and future, "holds the Earth in common" [@oxford]. They urge older generations to wake up from their profit-driven ignorance and comfort [@villasenor]. But how does age really affect self-reported concern about climate change? To answer this question, the following pages present a secondary analysis of survey data collected by Environics, commissioned by the City of Toronto, on residents' perceptions of climate change. Using the statistical programming language R, this study performs logistic regression to determine whether a correlation exists between two key variables: age and self-reported level of concern about climate change on a local and global scale. Informed by the youth climate movement, the hypothesis is that younger generations are more inclined to self-describe as concerned about climate change on both scales.

Drawing upon the activism that advocates for an intersectional approach to climate action, a multivariate regression analysis is also performed to include gender, education, income, and region. Intersectionality, as defined by Kimberle Crenshaw and described by Nina Lykke, is the idea that discrimination and systemic inequalities overlap – or intersect – across factors like gender, race, and class, and do not exist in isolation [@lykke]. Intersectional environmentalism, therefore, promotes climate action that recognizes how climate change disproportionately affects the already disadvantaged [@madson]. In this case, the hypothesis is that participants at the margins will have a greater likelihood of reporting a greater level of concern about climate change.

The results provide valuable insight. The models indicate that when it comes to whether a Toronto resident self-describes as concerned or not concerned about climate change as it affects Toronto and the Greater Toronto Area, gender, education, and income have a statistically significant effect. For reports on concern about climate change as it affects the rest of the world, only gender and education are important. Regarding whether Torontonians claim to be informed about climate change or not, the most important factor is education, unsurprisingly, but also gender. As argued by Jackson and De Coninck, effective climate action, particularly in adaptation and resilience, depends on active participation at a local level [@jackson]. With this in mind, this paper advocates for the use of open data practices by local governments, including municipalities, to promote data-informed decision-making. For civic engagement and grassroots organizing, access to reliable data is vital. It allows organizers to direct outreach efforts and allocate precious resources efficiently. Due to the scale and complexity of the challenges posed by climate change, ensuring the public has access to quality data can prove immensely beneficial. As poignantly illustrated by Kieran Healy, the crowdsourced data and data analyses that continue to uncover the effects of COVID-19 demonstrate the power of collaboration and open data in addressing challenges of unforeseen scales [@healy]. Similarly, climate change - and effective climate action - have a lot to gain from public ingenuity fueled by open data.

The paper first presents an overview of the original dataset and the variables relevant to this study. It also provides critical commentary on the data collection methods and original survey design. In addition, it compares the surveyed sample to the provincial and national populations. Next, the Methodology section discusses the strengths and weaknesses of the implemented model, as well as the model that was considered as an alternative. It includes details on how the data were re-coded to suit the model and how non-response was addressed. Then, the Results section explains the baseline category for all models and an interpretation of the estimated coefficients from each analysis. Finally, the Discussion section is comprised of four subsections: Findings, Ethics & Implications, and Limitations & Future Work situating the study within a broader context.

```{r codeimportdata, include = FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Import data

# The nice way*:
# search_packages("climate") %>% 
#   list_package_resources() %>% 
#   filter(grepl("v1", name)) %>% 
#   get_resource()

# * commented out because there ia an issue with the original file extension (listed as .xls but behaves as .xlsx)
# contacted data owner on 29032021, issue persists as of 17042021 with error:
#   "filepath: /var/folders/ff/yckwhmhx69jbjdlphv588lv80000gn/T//RtmpAD7G0d/file14a6c14da63d9.xls"
#   "libxls error: Unable to open file"

# Here's a less elegant but functional workaround:
climate_perceptions_raw <- getURL("https://raw.githubusercontent.com/lalmaraz/toronto_climate_perceptions/main/inputs/data/csv_city-of-toronto-climate-perceptions-spss-v1.csv")
climate_perceptions_csv <- read.csv(text = climate_perceptions_raw)

# Save
write_csv(climate_perceptions_csv, here("outputs/data/climate_data.csv"))
```

# Data
This report, including all necessary data cleaning, analysis and visualization, was produced with the R statistical programming language (R Core Team 2020) in the R Markdown file format. It uses features from several packages. The dataset is accessed through the *opendatatoronto* package [@opendatatoronto]. The *tidyverse* package is used for data cleaning and manipulation [@tidyverse], as well as *janitor* [@janitor]. Tables and graphs are generated with *kableExtra* [@kableExtra], *extrafont* [@extrafont], *modelsummary* [@modelsummary], *patchwork* [@patchwork], and *ggplot2* [@ggplot2]. A map is created with *ggmap* [@ggmap], *osmdata* [@osmdata] and *sf* [@sf]. External data is accessed, imported, and exported with *RCurl* [@RCurl], *readxl* [@readxl], *httr* [@httr], *xml2* [@xml2], and *rvest* [@rvest]. The models are assessed with *performance* [@performance]. To facilitate a reproducible workflow, *here* [@here] is used to reference file locations, and *bookdown* [@bookdown] and *knitr* [@knitr] to format the report.

The dataset is publicly available through the Open Data Portal hosted by the City of Toronto, and as mentioned, it is accessed through the *opendatatoronto* package [@opendatatoronto]. The data was last refreshed on August 15, 2019, and is listed in the catalogue as “Climate Perception Study”, published by the Environment & Energy Division. The data, as the catalogue states, was collected by Environics Research, a private Canadian polling and market research firm based in Toronto. The firm was commissioned by the City of Toronto to conduct an online survey with a representative sample of Toronto residents [@opendataportal]. According to the source, the survey was designed to “capture perceptions about climate change” to “help the City establish goals, priorities and metrics that will measure success in [climate action]” [@opendataportal]. The catalogue includes a report with results from data analysis prepared by Environics under the section "More information". This report, titled "City-Wide Climate Perceptions Study" [@environics] is referenced throughout this paper to compare results.

The data is provided in two versions in the Open Data Portal, a raw version (V1) and a secondary version coded for quantitative analysis (V2). For the purposes of this paper, and to code the data to this paper’s specific aims, only V1 data is used.

The survey was conducted in 2018, between October 11 and 18, with a total of 404 adult residents of Toronto answering the online questionnaire [@opendataportal]. The study followed a purposive sampling method, having specific predetermined groups of interest, residents aged 18 years and over, and implementing screening questions at the beginning of the questionnaire. As the source claims, sample quotas were set by region, age, and gender, in keeping with federal 2016 Census data [@opendataportal], presumably following a proportional approach. The study description on the Open Data portal lists that the survey was conducted on a sample of convenience, indicating that the survey respondents were not randomly selected and invited to participate based on systematic methods, but rather based on proximity, availability, or ease. The specific strategies used to recruit participants are not explained. Additionally, the source does not disclose why sampling was limited to adult residents.

The survey collected a wide range of data, including demographic and general information like postal code, gender, age, language of preference, income level, education level, type of residence, and homeownership. In addition, the survey also asked respondents attitudinal and behavioural questions including their level of concern about climate change as it affects Toronto and the Greater Toronto Area, level of concern of climate change as it affects the rest of the world, level of information about climate change, perceived impacts of climate change in the community, awareness of climate action programs nationally and locally, and believes, attitudes, and behaviours towards climate change and climate action. The dataset spans 152 columns which include, in addition to the participants' direct responses, systemic thematic coding for open-ended questions and all possible responses for question loops. A preview of a few columns and rows from the original dataset as provided by the source can be found in Table \@ref(tab:tabdatapreview).

```{r tabdatapreview, echo = FALSE, message = FALSE, warning = FALSE}
# Create a table that shows a preview of the original dataset
climate_perceptions_csv %>% 
  slice(9:15) %>% 
  select(5, 8, 10, 11, 12, 13, 14) %>%
  knitr::kable(caption = "A preview of the Climate Perceptions Study dataset", booktabs = T) %>% 
  kable_styling(latex_options = c("scale_down"))
```

From the dataset, this paper focuses on the following: (1) postal code, a nominal variable; (2) age, measured at the ordinal level; (3) gender, nominal; (4) level of concern of the local impact of climate change, (5) level of concern of the global impact of climate change, (6) level of information about climate change, (7) education level, and (8) income level, all ordinal variables. Table \@ref(tab:tabfocusdata) shows a summary of key demographic features subdivided into a binary, "yes" or "no", for each of the three dependent variables: concerned about climate change as it affects Toronto and the Greater Toronto Area (local), concerned about climate change as it affects the rest of the world (global), and informed about climate change. A list of the questions used to collect this information, as included in the original questionnaire [@questionnaire], can be found in [Appendix 1](#appendix1).

The original report separated responses based on the participants' district of residence. Similarly, using the first three digits of respondents' postal code and corresponding approximate latitude and longitude, a map with the approximate locations of respondents can be found in Figure \@ref(fig:respmap).

```{r codeinterestvariables, echo = FALSE, message = FALSE, warning = FALSE}
# Select variables of interest
climate_perceptions <- 
  climate_perceptions_csv %>%
  select(QS1, QS2, HIDAGE1, Q1r1, Q1r2, Q2, QD5, QD6) %>%
  rename(postal_code = QS1, gender = QS2, age = HIDAGE1, concern_local = Q1r1, concern_global = Q1r2, information = Q2, education = QD5, income = QD6)

# Create binary options for outcomes: concern_local, concern_global, and information
concerned_local <- c("Extremely concerned", "Very concerned")
not_concerned_local <- c("Not very concerned", "Not at all concerned")

concerned_global <- c("Extremely concerned", "Very concerned")
not_concerned_global <- c("Not very concerned", "Not at all concerned")

informed <- c("Extremely informed", "Very informed")
not_informed <- c("Not very informed", "Not at all informed")

# Add column for each binary to dataframe
climate_perceptions <- climate_perceptions %>%
  mutate(concerned_local_binary = case_when(
    concern_local %in% concerned_local ~ TRUE,
    concern_local %in% not_concerned_local ~ FALSE
  ))

climate_perceptions <- climate_perceptions %>%
  mutate(concerned_global_binary = case_when(
    concern_global %in% concerned_global ~ TRUE,
    concern_global %in% not_concerned_global ~ FALSE  
  ))

climate_perceptions <- climate_perceptions %>%
  mutate(informed_binary = case_when(
    information %in% informed ~ TRUE,
    information %in% not_informed ~ FALSE 
  ))

# Save
write_csv(climate_perceptions, here("outputs/data/climate_data.csv"))
